---
title: "PHP 2514 HW 6"
author: "Blain Morin"
date: "May 8, 2018"
output: html_document
---

# Gelman, Hill Chapter 14 #3 

3. Three-level logistic regression: the folder rodents contains data on rodents in a
sample of New York City apartments.

(a) Build a varying intercept logistic regression model (varying over buildings) to
predict the presence of rodents (the variable rodent2 in the dataset) given
indicators for the ethnic groups (race) as well as other potentially relevant
predictors describing the apartment and building. Fit this model using lmer()
and interpret the coefficients at both levels.

(b) Now extend the model in (b) to allow variation across buildings within com-
munity district and then across community districts. Also include predictors

describing the community districts. Fit this model using lmer() and interpret
the coefficients at all levels.

(c) Compare the fit of the models in (a) and (b).


## (a)

First, we fit a varying intercept logistic model. The variables that we choose to examine are:

-Rotted / Broken Windows (extwin4_2)
-Missing or Worn Flooring (extflr5_2)
-Cracks in Walls (intcrack2)
-Holes in the flooring (inthole2)
-If there is a regular exterminator (regext)
-Total household income (totincom2)


```{r, message=FALSE}

library(readr)
library(lme4)
library(sjPlot)
rats = read.table("rodents.txt")

attach(rats)

### Change to factors
rats$rodent2 = as.factor(rodent2)
rats$extwin4_2 = as.factor(extwin4_2)
rats$extflr5_2 = as.factor(extflr5_2)
rats$intcrack2 = as.factor(intcrack2)
rats$inthole2 = as.factor(inthole2)
rats$old = as.factor(old)
rats$regext = as.factor(regext)
rats$housing = as.factor(housing)
rats$race = as.factor(race)




### Run Model
model.a = glmer(rodent2 ~ 
                  extflr5_2 +
                  intcrack2 +
                  inthole2 +
                  regext +
                  as.factor(race) +
                  (1 | bldg),
                family = binomial (link = "logit"), 
                control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 10000000)))


summary(model.a)
sjt.glmer(model.a)


```


## (b)

```{r}

rats$cd = as.factor(cd)

model.b = glmer(rodent2 ~ 
                  extflr5_2 +
                  intcrack2 +
                  inthole2 +
                  regext +
                  as.factor(race) +
                  (1 | cd/bldg) +
                  (ownhous_Mean + old_Mean | cd),
                family = binomial (link = "logit"),
                control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 10000000)))

summary(model.b)
sjt.glmer(model.b)

```

questions: within cd vs across cd?

## (c)

```{r}

anova(model.a, model.b)

```
